<% set j = d['javadoc-data.json|javadocs'] %>
<% set nested_subclasses = a['javadoc-data.json|javadocs'].filter_class().nested_subclasses %>

\section{Curves}

The abstract Curve and its subclasses represent one-dimensional functions, y = f(x).

The following classes implement curves.

\begin{fullwidth}
<% for indent, class_name in nested_subclasses(j, 'com.opengamma.math.curve.Curve') %>
<% set package_name, x, cn = class_name.rpartition(".") %>
<< indent * "$\hookrightarrow$ " >><< class_name >> \href{http://docs-static.opengamma.com/<< OG_VERSION >>/analytics/api/<< package_name >>.html#class-<<class_name>>}{(api)} \href{http://docs-static.opengamma.com/<< OG_VERSION >>/java/javadocs/<< class_name.replace(".","/") >>.html}{(javadoc)}
<% endfor -%>
\end{fullwidth}

Here is an example of creating a simple curve:
<< a['examples-jython/curve-construction/curve-construction.py|idio|jythoni|pyg|l'].output_text() >>

\section{Surfaces}

Classes of surfaces represent two-dimensional functions, z = f(x,y).

The following classes implement surfaces.

\begin{fullwidth}
<% for indent, class_name in nested_subclasses(j, 'com.opengamma.math.surface.Surface') %>
<% set package_name, x, cn = class_name.rpartition(".") %>
<< indent * "$\hookrightarrow$ " >><< class_name >> \href{http://docs-static.opengamma.com/<< OG_VERSION >>/analytics/api/<< package_name >>.html#class-<<class_name>>}{(api)} \href{http://docs-static.opengamma.com/<< OG_VERSION >>/java/javadocs/<< class_name.replace(".","/") >>.html}{(javadoc)}
<% endfor -%>
\end{fullwidth}

\section{Cubes}

Cubes represent generic functions in 3D, v = f(x,y,z).

The following classes implement cubes.

\begin{fullwidth}
<% for indent, class_name in nested_subclasses(j, 'com.opengamma.math.cube.Cube') %>
<% set package_name, x, cn = class_name.rpartition(".") %>
<< indent * "$\hookrightarrow$ " >><< class_name >> \href{http://docs-static.opengamma.com/<< OG_VERSION >>/analytics/api/<< package_name >>.html#class-<<class_name>>}{(api)} \href{http://docs-static.opengamma.com/<< OG_VERSION >>/java/javadocs/<< class_name.replace(".","/") >>.html}{(javadoc)}
<% endfor -%>
\end{fullwidth}

\section{Matrices}

The Matrix interface provides a basis for generic multi-dimensional arrays.

The following classes implement matrices.

\begin{fullwidth}

com.opengamma.math.matrix.Matrix \href{http://docs-static.opengamma.com/<< OG_VERSION >>/analytics/api/com.opengamma.math.matrix.Matrix.html#class-com.opengamma.math.matrix.Matrix}{(api)} \href{http://docs-static.opengamma.com/<< OG_VERSION >>/analytics/javadoc/com/opengamma/math/matrix/Matrix}{(javadoc)}
<% for impl_class_name in d['javadoc-data.json|javadocs']['packages']['com.opengamma.math.matrix']['classes']['Matrix'].implementers %>
<% for indent, class_name in nested_subclasses(j, impl_class_name) %>
<% set package_name, x, cn = class_name.rpartition(".") %>
<< indent * "$\hookrightarrow$ " >><< class_name >> \href{http://docs-static.opengamma.com/<< OG_VERSION >>/analytics/api/<< package_name >>.html#class-<<class_name>>}{(api)} \href{http://docs-static.opengamma.com/<< OG_VERSION >>/java/javadocs/<< class_name.replace(".","/") >>.html}{(javadoc)}
<% endfor -%>
<% endfor -%>

\end{fullwidth}

